---
title: "Intervention simulation"
author: "Jan Netík"
date: "21 3 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load EpiModel
suppressMessages(library(EpiModel))

control <- control.icm(type = "SIR", nsteps = 100, nsims = 5)

init <- init.icm(s.num = 1318998, i.num = 125, r.num = 2)

param <- param.icm(inf.prob = 0.1181667, act.rate = 10, rec.rate = 1/20) 
    # a.rate = (10.5/365)/1000, ds.rate = (7/365)/1000, di.rate = (3.5/365)/1000, 
    # dr.rate = (7/365)/1000)

sim <- icm(param, init, control)

sim

plot(sim)

# the incidence (count of new cases per day).
plot(sim, y = "si.flow", mean.col = "red", qnts.col = "red")


```
```{r}
library(distcrete)
si_distr_lognorm <- distcrete("lnorm", interval = 1, 2.02, 2.78)
# $r(2000) %>% extract(.<20)

# freq_lognorm <- table(si_distr_lognorm) %>% as.vector
# si_distr_lognorm <- c(0, freq_lognorm/sum(freq_lognorm))
```

## R_O according to simulation and previously known SI distr.

```{r}
library(incidence)
library(earlyR)
incidence_counts <- as.data.frame(sim, out="mean") %>%
  select(time, si.flow)
incident_case_dates <- incidence_counts %>%
                        uncount(si.flow) %>%
                        pull(time) 
incidence_all <- incident_case_dates %>%
                    incidence(.)
peak_of_epidemic_curve <- find_peak(incidence_all)
incident_case_dates_growth_phase <- incidence_counts %>%
                            filter(time <= peak_of_epidemic_curve) %>%
                            select(time, si.flow) %>%
                            uncount(si.flow) %>%
                            pull(time)
incidence_growth_phase <- incident_case_dates_growth_phase %>%
                            incidence(., last_date=peak_of_epidemic_curve)
si <- si_distr_lognorm
res <- get_R(incidence_growth_phase, si = si)
plot(res, "R")
```

```{r}
run_sir_sim <- function(inf_prob, act_rate, pop_size = 1000, 
    i_num = 3, n_steps = 365, n_sims = 10, si_mean = 7.5, si_sd = 3.4) {
    
    # set up simulation parameters
    param <- param.icm(inf.prob = inf_prob, act.rate = act_rate, 
        rec.rate = 1/20, a.rate = (10.5/365)/1000, ds.rate = (7/365)/1000, 
        di.rate = (3.5/365)/1000, dr.rate = (7/365)/1000)
    init <- init.icm(s.num = pop_size - i_num, i.num = i_num, 
        r.num = 0)
    control <- control.icm(type = "SIR", nsteps = n_steps, nsims = n_sims)
    
    # run the simulation
    sim <- icm(param, init, control)
    
    # collect the relevant results in a data frame
    incidence_rates <- as.data.frame(sim, out = "mean") %>% select(time, 
        si.flow, i.num) %>% mutate(act_rate = act_rate, inf_prob = inf_prob, 
        total_cases = sum(si.flow), max_prev = max(i.num, na.rm = TRUE))
    
    # use the data frame of results to create an incidence()
    # object
    local_case_dates <- incidence_rates %>% filter(time <= 300, 
        act.rate == act_rate, inf.prob == inf_prob) %>% select(time, 
        si.flow) %>% uncount(si.flow) %>% pull(time)
    
    if (length(local_case_dates) > 0) {
        
        local_cases <- local_case_dates %>% incidence(.)
        
        # find the incidence peak from the incidence object
        peaky_blinder <- find_peak(local_cases)
        
        # recreate the incidence object using data only up to the
        # peak
        local_growth_phase_case_dates <- incidence_rates %>% 
            filter(time <= peaky_blinder) %>% select(time, si.flow) %>% 
            uncount(si.flow) %>% pull(time)
        
        local_growth_phase_cases <- local_growth_phase_case_dates %>% 
            incidence(., last_date = peaky_blinder)
        
        # get a MLE estimate of the basic reproduction number, R0
        res <- get_R(local_growth_phase_cases, si_mean = si_mean, 
            si_sd = si_sd)
        
        # add that as a column to the data frame of results
        incidence_rates <- incidence_rates %>% mutate(mle_R0 = res$R_ml)
    } else {
        # can't calculate R0 so just set to NA
        incidence_rates <- incidence_rates %>% mutate(mle_R0 = NA)
    }
    # return the data frame (which has just one row)
    return(incidence_rates)
}  # end function definition
```



Now we can run our simulation function in a loop, stepping through the levels of our two “public health intervention” parameters, and collect the results in an omnibus data frame.



```{r}
# set up an empty data frame to which to append results from
# each simulation
sims_incidence_rates <- tibble(time = integer(0), si.flow = numeric(0), 
    i.num = numeric(0), act_rate = numeric(0), inf_prob = numeric(0), 
    total_cases = numeric(0), max_prev = numeric(0), mle_R0 = numeric(0))

# the parameters to step through
act.rates <- c(20, 10, 2)
inf.probs <- c(0.05, 0.1, 0.15)

# loop through the parameter space
for (act.rate in act.rates) {
    for (inf.prob in inf.probs) {
        sims_incidence_rates <- sims_incidence_rates %>% bind_rows(run_sir_sim(inf.prob, 
            act.rate))
    }
}
```


OK, now we can visualise the results. First let’s look at the daily incidence (number of new cases per day):



```{r}
# create facet columns as descending ordered factors
sims_incidence_rates <- sims_incidence_rates %>% mutate(act_rate_facet_label = paste(act_rate, 
    "exposures per day"), inf_prob_facet_label = paste("Probability of infection\nat each exposure:", 
    inf_prob)) %>% arrange(desc(act_rate)) %>% mutate_at(vars(act_rate_facet_label), 
    funs(factor(., levels = unique(.)))) %>% arrange(desc(inf_prob)) %>% 
    mutate_at(vars(inf_prob_facet_label), funs(factor(., levels = unique(.)))) %>% 
    arrange(desc(act_rate), desc(inf_prob), time)

# add annotation text for each facet
sims_incidence_rates_facet_annotations <- sims_incidence_rates %>% 
    mutate(label = paste("R0 =", format(mle_R0, digits = 3), 
        "\n", round(100 * total_cases/1000, digits = 0), "% of population infected")) %>% 
    select(inf_prob_facet_label, act_rate_facet_label, label) %>% 
    distinct()

sims_incidence_rates %>% filter(time <= 365) %>% ggplot(aes(x = time, 
    y = si.flow)) + geom_line(colour = "blue", size = 1.5) + 
    facet_grid(inf_prob_facet_label ~ act_rate_facet_label) + 
    geom_text(data = sims_incidence_rates_facet_annotations, 
        mapping = aes(x = 50, y = 0.8 * max(sims_incidence_rates$si.flow, 
            na.rm = TRUE), label = label), parse = FALSE, hjust = 0, 
        vjust = 0, size = 3) + labs(x = "Days since start of epidemic", 
    y = "New cases per day", title = "Modelling of new cases of COVID-19 per day: incidence rate", 
    subtitle = paste("with varying levels of social mixing (exposures per day)", 
        "and probabilities of infection at each exposure")) + 
    theme(legend.position = "top", strip.text = element_text(size = 14))
```
```{r, fig.width=10, fig.height=8, eval=TRUE}
sims_incidence_rates %>%
  filter(time <= 365) %>%
  ggplot(aes(x=time, y=i.num)) +
  geom_line(colour="orange", size=1.5) +
  facet_grid(inf_prob_facet_label ~ act_rate_facet_label) +
  geom_text(data=sims_incidence_rates_facet_annotations,
            mapping=aes(x = 50, 
                        y = 0.8*max(sims_incidence_rates$i.num, na.rm=TRUE),
                        label = label), parse=FALSE,
            hjust   = 0, vjust   = 0, size=3) +
  labs(x="Days since start of epidemic", y="Prevalent (current number of active) cases",
       title="Modelling of new cases of COVID-19 per day: case prevalence",
       subtitle=paste("with varying levels of social mixing (exposures per day)",
                      "and probabilities of infection at each exposure")) +
  theme(legend.position = "top", 
          strip.text = element_text(size=12)) 
```

An alternative visualisation puts all the incidence curves on one panel:

```{r, fig.width=10, fig.height=8, eval=TRUE}
sims_incidence_rates %>%
  filter(time <= 365) %>%
  ggplot(aes(x=time, y=i.num, fill=act_rate_facet_label)) +
  geom_area(stat="identity", alpha=0.6) +
  facet_grid(inf_prob_facet_label ~ .) +
  labs(x="Days since start of epidemic", y="Prevalent (current number of active) cases",
       title="Modelling of prevalence of COVID-19 cases per day",
       subtitle=paste("with varying levels of social mixing (exposures per day)",
                      "and probabilities of infection at each exposure")) +
  theme_minimal() +
  theme(legend.position = "top", 
        legend.title = element_blank(),
          strip.text = element_text(size=12)) +
  scale_fill_brewer(type="seq", palette="Oranges")
```


```{r}
library(ggridges)

# create one column for both intervention parameters
sims_incidence_rates <- sims_incidence_rates %>% mutate(intervention_level_label = paste(act_rate, 
    "exp/day,", inf_prob * 100, "% inf risk/exp")) %>% arrange(max_prev, 
    time) %>% mutate_at(vars(intervention_level_label), funs(factor(., 
    levels = unique(.), ordered = TRUE)))

sims_incidence_rates %>% filter(time <= 365) %>% ggplot(aes(x = time, 
    y = intervention_level_label, height = i.num, fill = intervention_level_label)) + 
    geom_density_ridges(stat = "identity", show.legend = FALSE) + 
    labs(x = "Days since start of epidemic", y = "Prevalent (current number of active) cases", 
        title = "Modelling of COVID-19 transmission in 1,000 simulated people", 
        subtitle = paste("with varying levels of social mixing (exposures per day)", 
            "and risk of infection at each exposure,\n", "ordered by descending maximum number of prevalent cases per day")) + 
    theme_minimal() + theme(legend.position = "top", strip.text = element_text(size = 12)) + 
    scale_fill_brewer(type = "seq", palette = "Oranges")
```

